Design and Implementation of Probabilistic Programming Language Anglican

### 2017-03-14 paper/paper.tex
small further edits for consistency of use of terminology
JW in NEU

\newtoggle{full}
\input{toggle-ifl}

@@

\author{Jan-Willem van de Meent}
\affiliation{\institution{University of Oxford}}
\email{jwvdm@robots.ox.ac.uk}

@@

A probabilistic program, or query, mostly runs deterministic code.
Aside from the special forms \texttt{sample} and \texttt{observe},
which are probabilistic in nature, Anglican can be implemented as a
regular programming language. At \texttt{sample} and
\texttt{observe} forms, normal deterministic execution is 
interrupted and Anglican programs must allow the inference algorithm
to step in, recording information and affecting control flow. We refer these points in the execution as checkpoints. Handling of 
checkpoints can be implemented through coroutines/cooperative multitasking, and parallel execution/preemptive multitasking, as well as through explicit maintenance of program continuations. Anglican follows the latter option.

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved by returning a record that represents the state of program execution at the checkpoint, rather than invoking a continuation. The Clojure record used to interrupt execution at 
\texttt{sample} and \texttt{observe} forms are \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} and \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe} respectively. These records contain a unique identifier (which we will discuss in detail in section \ref{sec:addressing}), the arguments to the special form, the continuation, and the state.
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) 'cont)
(->sample 'S1 dist cont $state)
=> (cps-of-expression '(observe dist v) 'cont)
(->observe 'O1 dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the
CPS-transformed expressions are constructors for Clojure
records, and they take values of their fields as arguments.
Calling the continuation resumes the computation.

>>>>

\newtoggle{full}
\input{toggle}

@@

\author{Jan-Willem van de Meent}
\affiliation{\institution{Northeastern University}}
\email{j.vandemeent@northeastern.edu}

@@

A probabilistic program, or query, mostly runs deterministic code,
except for \texttt{sample} and \texttt{observe} forms, which are
probabilistic in nature. For this reason, Anglican can be implemented
as a regular programming language, except for the handling of these
two special forms, which we refer to as checkpoints. At these
checkpoints, Anglican programs must allow the inference algorithm to
step in, recording information and affecting control flow. This can
be implemented through coroutines/cooperative multitasking, and
parallel execution/preemptive multitasking, as well as through
explicit maintenance of program continuations at checkpoints.
Anglican follows the latter option.

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved by  returning a \textbf{checkpoint} instead of
invoking a continuation in a CPS-transformed program.  
Checkpoints are represented by values of the
corresponding record types
(\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} or
\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe})
in Clojure. The types contain fields specific to each form, as well as the
continuation; calling the continuation resumes the computation.
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) 'cont)
(->sample dist cont $state)
=> (cps-of-expression '(observe dist v) 'cont)
(->observe dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the
CPS-transformed expressions are constructors for Clojure
records, and they take values of their fields as arguments.
